### 2021-12-23 old_notes.tex

\section{Noted dynamic events}
We see numerous propagating disurbances in the \halpha{} linewidth.

ALMA shows lots of oscillations, difficult to say if these are impulsive or not (ALMA is both low resolution and radiatively thick).

Many brightenings in XRT. A few can be connected through to IBIS.
\begin{description}
    {\item[NuSTAR]-potential events}
    \begin{itemize}
        {\item 19:04}
        {\item 19:30}
    \end{itemize}
    {\item[IBIS]-potential events}
        \begin{itemize}
            {\item 14:53-14:58: propagating linewidth enhancement (-75,-40), (scan 158 )}
            {\item 15:15-15:18: likely bogus, I think data is too poor to tell}
            {\item 15:589-16:02: inflow to center of region (-55,-55).  See $d\lambda$ = 0.0320$\rightarrow$0.069}
            {\item 16:20-16:22: propagating wavetrain outward from center around (-50,-40).  See running difference $d\lambda$ = -0.089, -0.075,0.057, 0.069.  Shows up especially well blue-wing intensity images at $d\lambda = -0.1, -0.089\unit{nm}$.  May start as early as 16:10?  Go from 16:10-16:30}
        \end{itemize}
    {\item[XRT] blah}
    {\item[ALMA] blah}
    {\item[IRIS] blah}
\end{description}